### 2019-02-21 concepts.tex
fix idemopotent
discuss deflaker and elaborate on nondeterminism vs. flaky tests

@@

Vertical nondeterminism, in contrast, is demonstrated by a single trace
$t$, not a pair of traces.  To define vertical nondeterminism with
respect to idempotency, we can
define $\mathcal{I} \subseteq A$, the set of \emph{idempotent}
actions.  A trace $t s_0
\xrightarrow[]{\alpha_0} s_1 \xrightarrow[]{\alpha_1} \ldots
\xrightarrow[]{\alpha_{n-1}} s_n$ shows \emph{vertical nondeterminism
  with respect to idempotent operations} if $\exists i < n$ such that
$\alpha_i \in \mathcal{I} \wedge \alpha_i = \alpha_{i+1}$ and $v(s_i)
\neq v(s_{i+2})$.

>>>>

\subsection{Nondeterminism vs. Flakiness}

A key concept in this paper is that flaky tests (tests that sometimes
pass and sometimes fail, recall) exist due to underlying
nondeterministic behavior.  Our interest is in identifying such
nondeterministic behavior, not in addressing flakiness \emph{per se}.
Of course, the concepts are closely related.  We make ``did not behave
nondeterministically'' an additional \emph{property} that a test can
fail.  Due to the nature of (horizontal) nondeterminism, such failure
is almost always going to be inherently flaky --- sometimes a test
will behave nondeterministically, and sometimes it will not.  But the
idea is a separate one from flakiness; in order to show horizontal
nondeterminism, we need only see a change in values in a test, not in
the disposition (failed/passed) of the test.  It may be that, without
the additional property of determinism, the change cannot actually
cause failure, and thus flakiness.  The real relationship between our
nondeterminism and flakiness is that tests (automatically generated or
manually constructued) that rely on code that can be shown to exhibit
horizontal nondeterminism contain a potential source of flakiness.  If
the code behaves nondeterministically, and the values that are
observed to change can flow to influence the outcome of the test,
flakiness may well result.

Vertical nondeterminism, in fact, is potentially \emph{completely} unrelated
to flakiness, in that some bugs (improper handling of hardware errors
in a file system, for example) produce \emph{reliably}
``nondeterministic'' (from a software library users' point of view)
behavior.  Vertical nondeterminism may rely on ``real'' nondeterminism
that could produce a flaky test, but it does not have to do so.

@@

Vertical nondeterminism, in contrast, is demonstrated by a single trace
$t$, not a pair of traces.  To define vertical nondeterminism with
respect to idempotency, we can
define $\mathcal{I} \subseteq A$, the set of \emph{idempotent}
actions.  A trace $t s_0
\xrightarrow[]{\alpha_0} s_1 \xrightarrow[]{\alpha_1} \ldots
\xrightarrow[]{\alpha_{n-1}} s_n$ shows \emph{vertical nondeterminism
  with respect to idempotent operations} if $\exists i < n$ such that
$\alpha_i \in \mathcal{I} \wedge \alpha_i = \alpha_{i+1}$ and $v(s_{i+1})
\neq v(s_{i+2})$.

### 2019-02-21 experiments.tex
discuss deflaker and elaborate on nondeterminism vs. flaky tests

>>>>

These results, of course, depend on the parameters of the check for
nondeterminism.  If we increase the delay to 0.01 seconds and the
number of replays of a test used to check for nondeterminism, we
almost double the number of nondeterministic tests discovered.  Both
visible value and final state approaches find 40 nondeterministic
tests, though the overlap of tests thus detected is not quite perfect
--- each method detects two tests the other does not.  The difference
in overhead, interestingly, is very similar:  0.48\%; however, under
these parameters, visible value nondeterminism is actually cheaper on average
than final state nondeterminism, due to early termination of the
additional replays, when nondeterminism is detected (the difference is
again significant by Wilcoxon test, with $p < 1.44 \times 10^{-10}$.

In practice, visible value nondeterminism and final state
nondeterminism will often be very similar in both ability to detect
nondeterministic behavior and overhead.  However, under unusual
conditions, the behavior of the approaches can be very different.  The
additional overhead of visible value nondeterminism is usually low
because re-executing a test, not comparing state values for equality, is
the primary cost in nondeterminism detection; however, if there are a
very large number of state components, or some state components have
an expensive equality check (e.g., recursive structures with cycle
detection or depth limits), the overhead can grow, proportional to the
length of tests under consideration.  Similarly, final state
nondeterminism in a setting such as {\tt redis-py}, where divergences
in behavior tend to propagate to other state components, or at least
persist until termination of a test, detects nondeterminism quite
effectively.  However, in a setting where changes in behavior can
easily by overwritten by future test behavior, and do not causally
influence future computation, final state nondeterminism may be almost useless.

### 2019-02-21 introduction.tex
discuss deflaker and elaborate on nondeterminism vs. flaky tests

This paper proposes (1) a number of formal definitions of types of
nondeterminism (\emph{horizontal} and \emph{vertical}) and (2) an implementation, based on these definitions, for detecting and debugging
nondeterminism in property-based testing.  The implementation is based
on an approach where (3) horizontal determinism is considered as a
kind of \emph{reflexive differential testing} and (4) vertical
determinism is specialized to the common case of \emph{failure determinsim}.
We further propose modifications to the widely used delta-debugging algorithm in order to better handle
nondeterminism as a test property

This enables a number of useful results:
 the automated production of
nondeterminism-free regression tests even for libraries with some
sources of nondeterminism, the documentation of
nondeterminism, and, most importantly, detection and effective debugging for
unexpected nondeterminism that is, itself, a bug.  We use a set
of case studies of real-world, widely-used, Python libraries to demonstrate the
utility of our ideas.  Failure determinism directly
produces a more than 5\% improvement in mutation score for an already highly
effective automated test generation tool for a file system, with no
additional specification or test design burden.

>>>>

This paper proposes (1) a number of formal definitions of types of
nondeterminism (\emph{horizontal} and \emph{vertical}) and (2) an implementation, based on these definitions, for detecting and debugging
nondeterminism in property-based testing.  The implementation is based
on an approach where (3) horizontal determinism is considered as a
kind of \emph{reflexive differential testing} and (4) vertical
determinism is specialized to the common case of \emph{failure determinsim}.
We further propose modifications to the widely used delta-debugging algorithm in order to better handle
nondeterminism as a test property.

The underlying idea in this paper is that most approaches to
nondeterminism/flaky test detection rely either on 1) identifying
potential sources of nondeterminism such as test inter-dependence
\cite{LamZE2015} or certain code smells \cite{palomba2017does}  or 2)
used a heuristic approach, such as that taken by DeFlaker
\cite{bell2018d}.  In the real world, people usually detect flaky
tests by the most obvious heuristic of all:  they observe a test both
fail and pass for the same code base.  In a sense, even DeFlaker
relies on this method, but avoids having to observe a flaky test actually
passing, by inferring that if a test was previously passing, and
executes no changed code, its failure after a code change \emph{must}
be due to flakiness/nondeterminism.  Our approach to the problem is
essentially different, and largely orthogonal; we concentrate on the detection and avoidance of
nondeterminism/flakiness using \emph{automated test generation},
and make it possible for a developer or test engineer to detect when
important values generated in automated testing differ without a
change in the test itself or the code under test.  While some of these
changes may be harmless, and unable to propagate to cause test
failures, they are potential sources of flaky behavior.  To our
knowledge, this is the first dynamic approach to the problem that
allows detection of (potential) flakiness, based on real divergence in
behavior, without having to observe a test actually fail due to nondeterminism.

This enables a number of useful results:
 the automated production of
nondeterminism-free regression tests even for libraries with some
sources of nondeterminism, the documentation of
nondeterminism, and, most importantly, detection and effective debugging for
unexpected nondeterminism that is, itself, a bug.  Moreover, by
identifying (using automated testing) library behaviors that are
nondeterministic, our approach should make it much easier to flag suspect
statements in human-written flaky unit tests.

We use a set
of case studies of real-world, widely-used, Python libraries to demonstrate the
utility of our ideas.  Failure determinism directly
produces a more than 5\% improvement in mutation score for an already highly
effective automated test generation tool for a file system, with no
additional specification or test design burden.

### 2019-02-21 relatedwork.tex
discuss deflaker and elaborate on nondeterminism vs. flaky tests

The problem of test nondeterminism is closely related to the (extremely important in industrial settings) issue of flaky tests \cite{miccoflaky, luo2014empirical,palomba2017does,listfieldtestanalysis}.  How to handle flaky tests in practice (when they cannot be avoided) is a major issue in Google-scale continuous testing \cite{memon2017taming}, and, as Memon et al. describe, the problem of flaky tests influences the general design of Google test automation.
Previous work on flaky tests has either focused on test inter-dependence as a cause of flaky behavior \cite{LamZE2015}, or provided large-scale empirical examination of tests from one large open source project (Apache) \cite{luo2014empirical,palomba2017does}.  Palomba and Zaidman \cite{palomba2017does} investigated the relationship between test code smells and flaky tests.   The present paper, rather than focusing on flaky tests per se, investigates new tools for handling nondeterminism in property-based test generation.  From our point of view, flaky behavior is simply a special case of nondeterminism, where the nondeterminism is sufficient to cause the test to have different pass/fail results at times.  Our approach is, in a sense (especially the delta-debugging modifications) in line with Harman and O'Hearn's proposal to simply accept that ``All Tests Are Flaky'' \cite{StartupstoScaleups}, and work with probabilistically failing tests.

>>>>

The problem of test nondeterminism is closely related to the (extremely important in industrial settings) issue of flaky tests \cite{miccoflaky, luo2014empirical,palomba2017does,listfieldtestanalysis}.  How to handle flaky tests in practice (when they cannot be avoided) is a major issue in Google-scale continuous testing \cite{memon2017taming}, and, as Memon et al. describe, the problem of flaky tests influences the general design of Google test automation.
Previous work on flaky tests has either focused on test inter-dependence as a cause of flaky behavior \cite{LamZE2015}, or provided large-scale empirical examination of tests from one large open source project (Apache) \cite{luo2014empirical,palomba2017does}.  Palomba and Zaidman \cite{palomba2017does} investigated the relationship between test code smells and flaky tests.

Bell et al. proposed DeFlaker \cite{bell2018d}, which makes flaky tests much easier to detect by relying on the observation that if a test fails, and does not cover any changed code then (1) presumably it was a passing test in the past, or the tests would not be ``green'' before the latest code change so (2) the test is likely flaky.  Comparing our approach with that of DeFlaker is difficult; DeFlaker applies only in the context of code \emph{changes} and is essentially a heuristic that identifies certain test failures as due to nondeterminism.  Our approach is primarily focused on automatically providing a more extensive \emph{specification} in automated test generation, where a test ``failing'' is itself a sign of nondeterministic behavior, that could give rise to flaky behavior.

The present paper, rather than focusing on flaky tests as such, therefore, investigates methods for handling nondeterminism in property-based test generation.  Again, from our point of view, flaky behavior is simply a special case of nondeterminism, where the nondeterminism is sufficient to cause the test to have different pass/fail results at times.  Our approach is, in a sense (especially the delta-debugging modifications) in line with Harman and O'Hearn's proposal to simply accept that ``All Tests Are Flaky'' \cite{StartupstoScaleups}, and work with probabilistically failing tests.
